Fig2: Candidate genes

### 2020-05-20 analysis/figure-2.Rmd
updated fig 2 and suppl fig 2

## table 1 summary candidate genes
table1 <- left_join(candidateDEGS, parentalcaregenes) %>%
  select(gene, bldg_lay:n5_n9, literature, GO, NCBI) %>%
  mutate(literature = if_else(is.na(literature), " ", "X"),
         GO = if_else(is.na(GO), " ", "X"))
table1[is.na(table1)] <- " " # replace NA with blank space so it's pretty
head(table1)

```

@@

```{r candidatevsds}
# load `candidatevsd` with `source("../R/wrangledata.R")`
candidatevsd <- read_csv("../results/03_candidatevsd.csv") %>% 
  select(-X1) %>%
  filter(treatment %in% charlevels)
tail(candidatevsd)

@@

```{r correlations}
hyp1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
pit1 <- makecorrdf("female", "pituitary", curleychampagnegenes)  
gon1 <- makecorrdf("female", "gonad", curleychampagnegenes)

@@

forlegend <- plotcorrplot(gon4, NULL) + theme(legend.position = "bottom") 
mylegend <- get_legend(forlegend)

@@

a <- png::readPNG("../figures/images/fig_fig2a.png")
a <- ggdraw() +  draw_image(a, scale = 1)

c1 <- scattercorrelations(FH, FH$DRD4, FH$HTR2C, "#969696", "DRD4", "HTR2C") + labs(subtitle = "female", title =  "Top correlation")  
c3 <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = NULL) 
c5 <- scattercorrelations(FG, FG$ESR1, FG$PGR, "#969696", "ESR1", "PGR")   + labs(subtitle = NULL)

c2 <- scattercorrelations(MH, MH$DRD4, MH$HTR2C, "#525252", "DRD4", "HTR2C")  + labs(subtitle = "male", title = "", y = NULL)  
c4 <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = NULL, y = NULL)  
c6 <- scattercorrelations(MG, MG$ESR1, MG$PGR, "#525252", "ESR1", "PGR")  + labs(subtitle = NULL, y = NULL)

d1 <- candidateboxplot("hypothalamus", c("DRD4"), sexlevels) + labs(x = NULL, title = "Temporal expression pattern" )  
d2 <- candidateboxplot("hypothalamus", c("HTR2C"), sexlevels) + labs(x = NULL ) + theme(strip.text = element_blank())
d3 <- candidateboxplot("pituitary", c("AVPR1A"), sexlevels) + labs(x = NULL )  + theme( strip.text = element_blank())
d4 <- candidateboxplot("pituitary", c("CRHR1"), sexlevels) + labs(x = NULL  )  + theme(strip.text = element_blank())
d5 <- candidateboxplot("gonad", c("ESR1"), sexlevels) + labs(x = NULL) + theme(strip.text = element_blank())
d6 <- candidateboxplot("gonad", c("PGR"), sexlevels) + labs(x = NULL ) + theme(strip.text = element_blank())

#b <- plot_grid(b1,b2,b3, ncol = 1, rel_heights = c(1.1,1,1))
c <- plot_grid(c1,c2,c3,c4,c5,c6, ncol = 2, rel_heights = c(1.1,1,1), rel_widths = c(1.1,1))
d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))

bcd <- plot_grid(c,d, ncol = 2, rel_widths = c(1,1), labels = c("A", "B"), label_size = 8)

fig2 <- plot_grid(bcd, a, ncol = 1, rel_heights = c(1,0.1))
fig2

pdf(file="../figures/fig2-1.pdf", width=7.25, height=7.25)
plot(fig2)
dev.off()
```

>>>>

## table 1 summary candidate genes
table1 <- left_join(candidateDEGS, parentalcaregenes) %>%
  select(gene, bldg_lay:n5_n9, literature, GO, NCBI) %>%
  mutate(literature = if_else(is.na(literature), " ", "X"),
         GO = if_else(is.na(GO), " ", "X"))
head(table1)
```

@@

```{r candidatevsds}
# load `candidatevsd` with `source("../R/wrangledata.R")`
candidatevsd <- read_csv("../results/03_candidatevsd.csv") %>% 
  select(-X1) %>%
  filter(treatment %in% charlevelsnocontrol) %>%
  drop_na()
tail(candidatevsd)

@@

```{r correlations}

makecorrdf <- function(whichsex, whichtissue, whichgenes){
  corrrdf <- candidatevsd %>%
    filter(sex == whichsex, tissue == whichtissue,
           gene %in% whichgenes) %>%
    drop_na() %>%
    pivot_wider(names_from = gene, values_from = counts) %>%
    select(-sex, -tissue, -treatment, -samples) %>%
    correlate() 
  # print(head(corrrdf))

  corrrdflong <- corrrdf %>%
    pivot_longer(-rowname, names_to = "gene2", values_to = "corr") %>%
    rename("gene1" = "rowname") %>%
    arrange(desc(corr)) %>%
    mutate(tissue = whichtissue,
           sex = whichsex) %>%
    select(tissue, sex, gene1, gene2, corr)
  print("The top two correlations in this tissue and sex are:")
  print(corrrdflong[c(1,3),])

  return(corrrdf)
}

hyp1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
pit1 <- makecorrdf("female", "pituitary", curleychampagnegenes)  
gon1 <- makecorrdf("female", "gonad", curleychampagnegenes)

@@

forlegend <- plotcorrplot(gon4, NULL) + theme(legend.position = "bottom") + labs(color = "Correlation")
mylegend <- get_legend(forlegend)

@@

candidateboxplot <- function(whichtissue, whichgenes, whichsex){

  p <- candidatevsd %>%
    filter(tissue %in% whichtissue,
           gene %in% whichgenes,
           sex %in% whichsex) %>%
    mutate(treatment = factor(treatment, levels = charlevels)) %>%
    ggplot(aes(x = treatment, y = counts)) +
    geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
    geom_jitter(size = 0.25, aes(color = sex)) +
    #facet_wrap(~sex,  nrow = 1) +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme_B3() + 
    theme(legend.position = "none",
          axis.title.y = element_text(face = "italic"),
          axis.text.x = element_blank(),
          axis.ticks = element_blank(),
          axis.line.x = element_blank()) +
    labs(y = whichgenes,
         x = NULL) +
    geom_signif(comparisons = list(c( "control", "bldg"),
                                   c( "bldg", "lay"),
                                   c( "lay", "inc.d3"),
                                   c("inc.d3", "inc.d9"),
                                   c( "inc.d9", "inc.d17"),
                                   c( "inc.d17", "hatch"),
                                   c("hatch", "n5"),
                                   c( "n5", "n9")),  
                map_signif_level=TRUE,
                textsize = 1.5, family = 'Helvetica',
                vjust = 1.5, size = 0)

  return(p)
}

c1 <- scattercorrelations(FH, FH$CRHR2, "CRHR2", FH$HTR2C, "HTR2C", "#969696" ) +  labs( title =  "Hypothlamic expression", subtitle = "females")  
c3 <- scattercorrelations(FP, FP$AVPR1A, "AVPR1A", FP$CRHR1,  "CRHR1", "#969696")  + labs(title = "Pituitary expression", subtitle = "females") 
c5 <- scattercorrelations(FG, FG$CRHBP, "CRHBP",  FG$ESR2,  "ESR2", "#969696")   + labs(title = "Gonadal expression", subtitle = "females")

c2 <- scattercorrelations(MH, MH$DRD1, "DRD1", MH$HTR2C, "HTR2C", "#525252")  + labs(title = " ", subtitle = "males" )  
c4 <- scattercorrelations(MP, MP$AVPR1A,  "AVPR1A",  MP$CRHR1, "CRHR1", "#525252")  + labs(title = " ", subtitle = "males" ) 
c6 <- scattercorrelations(MG, MG$PGR, "PGR", MG$PRLR, "PRLR", "#525252")   + labs(title = " ", subtitle = "males" )

d1 <- candidateboxplot("hypothalamus", c("CRHR2"), "female") + labs(x = NULL, title = " ", subtitle = " " )  
d2 <- candidateboxplot("hypothalamus", c("HTR2C"), "female") + labs(x = NULL )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
d3 <- candidateboxplot("pituitary", c("AVPR1A"), "female") + labs(x = NULL , subtitle = " " ) 
d4 <- candidateboxplot("pituitary", c("CRHR1"), "female") + labs(x = NULL  )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
d5 <- candidateboxplot("gonad", c("CRHBP"), "female") + labs(x = NULL, subtitle = " " )
d6 <- candidateboxplot("gonad", c("ESR2"), "female") + labs(x = NULL ) + theme(axis.text.x = element_text(angle = 45, vjust = 1))

e1 <- candidateboxplot("hypothalamus", c("DRD1"), "male") + labs(x = NULL, title = " " , subtitle = " " )  
e2 <- candidateboxplot("hypothalamus", c("HTR2C"), "male") + labs(x = NULL )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
e3 <- candidateboxplot("pituitary", c("AVPR1A"), "male") + labs(x = NULL , subtitle = " " ) 
e4 <- candidateboxplot("pituitary", c("CRHR1"), "male") + labs(x = NULL  )  + theme(axis.text.x = element_text(angle = 45, vjust = 1))
e5 <- candidateboxplot("gonad", c("PGR"), "male") + labs(x = NULL, subtitle = " " ) 
e6 <- candidateboxplot("gonad", c("PRLR"), "male") + labs(x = NULL )   + theme(axis.text.x = element_text(angle = 45, vjust = 1))

#b <- plot_grid(b1,b2,b3, ncol = 1, rel_heights = c(1.1,1,1))
c123 <- plot_grid(c1,c3,c5, ncol = 1, rel_heights = c(1.1,1,1), labels = c("A", "B", "C"), label_size = 8)
c456 <- plot_grid(c2,c4,c6, ncol = 1, rel_heights = c(1.1,1,1))
d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))
e <- plot_grid(e1,e2,d3,e4,e5,e6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))

fig2 <- plot_grid(c123,d,c456,e, nrow  = 1)
fig2

#pdf(file="../figures/fig2-1.pdf", width=7.25, height=7.25)
#plot(fig2)
#dev.off()
```
